---
title: "Visualizing election predictions using ggplot2"
output: html_document
---

Here, ggplot2 is used to visualize US presidential election predictions using logistic regression on polling data.


First, the data is separated into a training set, containing data from 2004 and 2008 polls, and a test set, containing the data from 2012 polls.Then a logistic regression model is developed to forecast the 2012 US presidential election.

```{r}

library(ggplot2)
library(maps)
library(ggmap)
require(downloader)
statesMap = map_data("state")
```
We can now plot the states as using ggplot2

```{r, fig.height=6,fig.width=10}

ggplot(statesMap, aes(x = long, y = lat, group = group)) + geom_polygon(fill = 'white', color = "black") +
theme(axis.title.y = element_text(colour="grey20",size=15,angle=90,hjust=.5,vjust=1,face="plain"),axis.title.x = element_text(colour="grey20",size=15,angle=0,hjust=.5,vjust=1,face="plain"),axis.text.y = element_text(colour="grey20",size=15,angle=0,hjust=1,vjust=0,face="plain"),axis.text.x = element_text(colour="grey20",size=15,angle=60,hjust=.5,vjust=.5,face="plain"))
```

Now, let's develop a logistic regression model and map the prediction results.

```{r}
url<-"https://courses.edx.org/asset-v1:MITx+15.071x_2a+2T2015+type@asset+block/PollingImputed.csv"

download(url, dest="PollingImputed.csv") 

polling = read.csv("PollingImputed.csv")

# Subset data into training set and test set
Train = subset(polling, Year == 2004 | Year == 2008)
Test = subset(polling, Year == 2012)

mod2 = glm(Republican~SurveyUSA+DiffCount, data=Train, family="binomial")

TestPrediction = predict(mod2, newdata=Test, type="response")


```

TestPrediction gives the predicted probabilities for each state, but let's also create a vector of Republican/Democrat predictions

```{r}
TestPredictionBinary = as.numeric(TestPrediction > 0.5)

```

To use ggplot2, predictions and state labels have to be in a data frame.

```{r}

predictionDataFrame = data.frame(TestPrediction, TestPredictionBinary, Test$State)

```

Now, we need to merge "predictionDataFrame" with the map data "statesMap". Before doing so, we need to convert the Test.State variable to lowercase, so that it matches the region variable in statesMap.

```{r}
predictionDataFrame$region = tolower(predictionDataFrame$Test.State)
predictionMap = merge(statesMap, predictionDataFrame, by = "region")

```

we need to make sure the observations are in order so that the map is drawn properly.

```{r}
predictionMap = predictionMap[order(predictionMap$order),]

```

Now, we can plot the predictions.

```{r,fig.height=6,fig.width=10}
ggplot(predictionMap, aes(x = long, y = lat, group = group, fill = TestPredictionBinary))+ geom_polygon(color = "black") + scale_fill_gradient(low = "blue", high = "red", guide = "legend", breaks= c(0,1), labels = c("Democrat", "Republican"), name = "Prediction 2012")

```

We can play with the color, bourder thickness, line type and transparency.



<br/>

```{r,fig.height=6,fig.width=10}
ggplot(predictionMap, aes(x = long, y = lat, group = group, fill = TestPrediction))+ geom_polygon(color = "black", alpha=0.3) + scale_fill_gradient(low = "blue", high = "red", guide = "legend", breaks= c(0,1))
```

<br/>

We can also plot the probabilities instead of the binary predictions.
Using the deafult color:

```{r,fig.height=6,fig.width=10}
ggplot(predictionMap, aes(x = long, y = lat,
    group = group, fill = TestPrediction,name='Fish'))+ 
    geom_polygon(color = "black",lty=3)+
    scale_fill_continuous(name="Probability of\nRepublican")

```

